---
title: "Multiple Linear Regresssion"
author: "R. Gevorgyan"
date: "October 21, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data
Here, we employ the CPS1988 data frame collected in the March 1988
Current Population Survey (CPS) by the US Census Bureau and analyzed
by Bierens and Ginther (2001).

These are cross-section data on males aged 18 to 70 with positive
annual income greater than US$ 50 in 1992 who are not self-employed
or working without pay.

```{r echo=T, results='hide', message=FALSE, warning=FALSE}
library(AER)
```

```{r }

data("CPS1988")
summary(CPS1988)
```

```
* wage - the wage in dollars per week, 
* education and experience - measured in years
* ethnicity is a factor with levels Caucasian ("cauc") and African-American ("afam"). 
* smsa - indicating residence in a standard metropolitan statistical area (SMSA) 
* region - the region within the United States of America, and
* parttime - whether the individual works part-time.
```
Note that the CPS does not provide actual work experience. It is therefore customary to compute experience as ***age - education - 6***; this may be considered potential experience. This quantity may become negative.

The model of interest is

***log(wage) = $\beta$~1~  + $\beta$~2~ experience + $\beta$~3~ experience2 +$\beta$~4~ education +$\beta$~5~ ethnicity +$\varepsilon$ ***

This is a semilogarithmic model, which can be fitted in R using


```{r}
cps_lm <- lm(log(wage) ~ experience + I(experience^2) +education + ethnicity, data = CPS1988)
summary(cps_lm)
```

The summary reveals that all coefficients have the expected sign, and the corresponding variables are highly significant (not surprising in a sample as large as the present
one). Specifically, according to this specification, the return on education is
8.57% per year.

To illustrate the general procedure for model comparisons, we explicitly fit the model without ethnicity and then compare both models using anova()

```{r}
cps_noeth <- lm(log(wage) ~ experience + I(experience^2) + education, data = CPS1988)
anova(cps_noeth, cps_lm)
```

