---
title: "Homework 4"
author: "Levon Khachatryan"
date: "December 18, 2018"
output: html_document
header-includes:
  - \usepackage{color}
---
<br />
<br />
**Recall the Homework**
<br />
<br />
we should split the data ( bankloan ) into train and test , do all the works with the train dataset and check the results in the test data set.
<br />
<br />
So let's start with importing necessary data.
```{r}
rm(list = ls())
logitdata = read.csv("D:/MyProjects/Econometrics_Homeworks/Homework_4/sources/bankloan_cs_noweights.csv", header = TRUE)
head(logitdata)
```
Now let's split the dataset into train and test.
```{r}
smp_size <- floor(0.75 * nrow(logitdata))

## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(logitdata)), size = smp_size)

train <- logitdata[train_ind, ]
test <- logitdata[-train_ind, ]
```
Dimension of train and test datasets are following.
```{r}
dim(train)
dim(test)
```
I am attaching a data set to cover future misunderstandings.
```{r}
# attach(logitdata)
attach(train)
attach(test)
```
Constructing model.
```{r}
logitmodel <- glm(default ~  ., family=binomial(link="logit"), data=train)
LogL<-logLik(logitmodel)
```
If we will do summary(logitmodel) , we will obtain the following.<br /><br />
Null deviance: 1.4876e+03  on 1124  degrees of freedom<br />
Residual deviance: 6.5268e-09  on    0  degrees of freedom<br />
AIC: 2250<br />
<br />
Number of Fisher Scoring iterations: 25<br />
```{r}
logitmodel1 <- glm(default ~  ed, family=binomial(link="logit"), data=train)
LogL1<-logLik(logitmodel1)

logitmodel2 <- glm(default ~  1, family=binomial(link="logit"), data=train)
LogL2<-logLik(logitmodel2)
log(mean(default)) - log(1 - mean(train$default))

McF<-1-LogL/LogL2
McF

McF1<-1-LogL1/LogL2
McF1
# install.packages("rcompanion")
library(rcompanion)
nagelkerke(logitmodel)

default.prob = predict(logitmodel, type="response")
default.pred = rep(0, nrow(train))
default.pred[default.prob > .5] = 1
table(default.pred, logitmodel$data$default)

# install.packages("pROC")
library(pROC)
roc(logitmodel$data$default,default.pred)
plot.roc(logitmodel$data$default,default.pred)
```
```{r echo=FALSE}
```
## Making prediction for test dataset.
```{r}
default.prob = predict(logitmodel, type="response")
default.pred = rep(0, nrow(test))
default.pred[default.prob > .5] = 1
table(default.pred, logitmodel$data$default)
```